---
title: "Proposed SeqPlots paper figures - *C. elegans*"
author: "Przemyslaw Stempor"
date: "July 30, 2015"
output:
  ioslides_presentation:
    fig_retina: null
    smaller: yes
    transition: 0
widescreen: yes
---


<style>
    .title-slide {
        background-color: #CBE7A5; /* #EDE0CF; ; #CA9F9D*/
    }
    
    slide {
        background-color: #FFFFFF; /* #EDE0CF; ; #CA9F9D*/
    }
    article.smaller pre {
        font-size: 12px;
        line-height: 13px;
        letter-spacing: 0;
    }
    
    slides > slide > hgroup + article {
        margin-top: 10px;
    }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

## Data aquisition

Following tracks (each containing averaged signal for 2 replicates) are used:

- `HTZ1^JA00001__IL1^F^N2^L3_IL010_and_IL2^F^N2^L3_IL009__NORM^linear^1bp_averaged.bw`
- `H3K4me3^ab8580__rc03^F^Cfp1-GFP^L3_RC007_and_rc04^F^Cfp1-GFP^L3_RC008__NORM^linear^1bp_averaged.bw`
- `H3K36me3^HK00001__PK07^F^N2^L3_PK007_and_PK06^F^N2^L3_PK006__NORM^linear^1bp_averaged.bw`

## Loading data and libraries
```{r aquire1, message=FALSE}
require(seqplots)
require(magrittr)
devtools::load_all("/Users/przemol/code/seqplotsR")

tracks <- dir('tracks', pattern = 'bw', full.names = TRUE)
features <- dir('features', pattern = 'bed', full.names = TRUE)

top_limits <- list(c(-0.3, 2.5), c(0, 6.5), c(-0.3, 3.5), c(6, 12))
colvec <- list(
    NA, 
    c('blue', 'white', 'red'), 
    c('blue', 'white', 'red'), 
    c('white', '#66ccff', '#000066')
)

tracks
features
```

## Data process
```{r aquire2, message=FALSE}
m1 <- getPlotSetArray(
    tracks = tracks,
    features = grep('top', features, value = TRUE),
    refgenome = 'ce10',
    bin = 10L, xmin = 1000L, xmax = 1000L
)

m2 <- getPlotSetArray(
    tracks = grep('H3K4me3', tracks, value = TRUE, ignore.case = TRUE),
    features = grep('top|third|bot', features, value = TRUE),
    refgenome = 'ce10',
    bin = 10L, xmin = 1000L, xmax = 1000L
)


# Bottom figures c and d
ms <- MotifSetup()
ms$addBigWig(tracks[2])$addBigWig(tracks[3])$addBigWig(tracks[1])
ms$addMotif('CG', revcomp = FALSE, name = "CpG")

M <- getPlotSetArray(
    tracks = ms,
    features = features[5],
    refgenome = 'ce10',
    bin = 10L, xmin = 1000L, xmax = 1500L
)

```

## Function for plotting c) and e)
```{r plotTopBar, message=FALSE}
plotTopBar <- function(M, titles=LETTERS, limits=NULL) {
    f = 37
    layout(
        matrix(c(1:100), 2, 4, byrow = FALSE), respect = TRUE,
        heights = c(.3, rep(1,5)), widths = c(1)
    )
    pp <- function() {
        par(mgp = c(0, 2.5, 0))
        axis(
            1, at = c(min(M[[1]]$all_ind), 0, max(M[[1]]$all_ind)),
            labels = c('-1kb', 'TSS', '+1.5kb'), cex.axis = f / 12
        )
        par(mgp = c(0, 1, 0))
    }
    plt <- function(p, ylim = NULL, ...) {
        for (i in 1:p$npaires()) {
            plot(
                p[i], ylim = ylim, keepratio = TRUE, legend = FALSE, ln.h = 0,
                cex.axis = f, cex.lab = f, ln.v = TRUE, xlab = '',
                cex.main = .01, panel.first = pp(), xaxt = "n", col = 'black', ...
            )
        }
    }
    
    par(mgp = c(3, 0, 0))
    
    plot.new(); title(xlab = titles[[1]], col.main = 'red', cex.lab = 6)
    plt(unlist(M[1,1]), ylim = limits[[1]], main = titles[[1]])
    
    par(mgp = c(3, 0, 0))
    
    plot.new(); title(xlab = titles[[2]], col.main = 'red', cex.lab = 6)
    plt(unlist(M[1,2]), ylim = limits[[2]], main = titles[[2]])
    
    par(mgp = c(3, 0, 0))
    
    plot.new(); title(xlab = titles[[3]], col.main = 'red', cex.lab = 6)
    plt(unlist(M[1,3]), ylim = limits[[3]], main = titles[[3]])
    
    par(mgp = c(3, 0, 0))
    
    plot.new(); title(xlab = titles[[4]], col.main = 'red', cex.lab = 6)
    plt(unlist(M[1,4]), ylim = limits[[4]], main = titles[[4]])
}
```

# Figures

## [Plot A](fa2.pdf)

```{r lineplot1, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='svg', fig.align='center', fig.height=8.27, fig.width=11.7, out.height=450, out.width="auto"}
plotAverage(
    m1, labels = colnames(m1$as.array())  %>% strsplit('_')  %>% sapply('[[', 1), 
    legend_pos = 'topleft', legend_ext = TRUE, legend_ext_pos = 'topright',
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22,
    main = "Top expression quintile", xlab = "TSS", ylab = "Z-scored ChIP signal"
)
dev.print(pdf, file='fa2.pdf'); dev.print(png, file='fa2.png', height=8.27, width=11.7, res=72, units = "in");
```

## [Plot B](fb2.pdf)

```{r lineplot2, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='svg', fig.align='center', fig.height=8.27, fig.width=11.7, out.height=450, out.width="auto"}
plotAverage(
    m2[3:1], labels = c("TSS top 20%", "TSS middle 20%", "TSS bottom 20%"), 
    legend_pos = 'topleft', legend_ext = TRUE, legend_ext_pos = 'topright',
    cex.legend = 17, cex.main = 36, cex.lab = 28, cex.axis = 22,
    main = "H3K4me3 ChIP-seq signal", xlab = "TSS", ylab = "Z-scored ChIP signal"
)
dev.print(pdf, file='fb2.pdf'); dev.print(png, file='fb2.png', height=8.27, width=11.7, res=72, units = "in");
```

## [Plot C](fc.pdf)
```{r topbar1, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='svg', fig.align='center', fig.height=2*4, fig.width=6*4, out.height='300', out.width="auto"}
plotTopBar(
    M, M$as.array()  %>% colnames %>% strsplit('_') %>% sapply('[[', 1),
    limits = top_limits
)
dev.print(pdf, file='fc.pdf'); dev.print(png, file='fc.png', height=2*4, width=6*4, res=72, units = "in");
```

## [Plot D](fd.pdf)
```{r heatmap1, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
cls <- plotHeatmap(
    M, labels = M$as.array()  %>% colnames %>% strsplit('_') %>% sapply('[[', 1),
    raster = TRUE, include = c(T,F,F,F), sortrows = 'decreasing', clusters = 3, 
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22, colvec = colvec,
    o_min = c(-1, -2, -1, 0), o_max = c(3, 8, 4, 16)
)

dev.print(pdf, file='fd.pdf'); dev.print(png, file='fd.png', height=8.27, width=11.7, res=72, units = "in");
```

## [Plot E](fe.pdf)
```{r topbar2, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='svg', fig.align='center', fig.height=2*4, fig.width=6*4, out.height='300', out.width="auto"}
export.bed(import.bed(features[5])[cls$ClusterID==1], 'flt.bed')
MF <- getPlotSetArray(
    tracks = ms,
    features = 'flt.bed',
    refgenome = 'ce10',
    bin = 10L, xmin = 1000L, xmax = 1500L
)
plotTopBar( 
    MF, MF$as.array()  %>% colnames %>% strsplit('_') %>% sapply('[[', 1),
    limits = top_limits
)
dev.print(pdf, file='fe.pdf'); dev.print(png, file='fe.png', height=2*4, width=6*4, res=72, units = "in");
```

## [Plot F](ff.pdf)
```{r heatmap2, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
plotHeatmap(
    MF, labels = MF$as.array()  %>% colnames %>% strsplit('_') %>% sapply('[[', 1),
    raster = TRUE, include = c(F,T,T,F), sortrows = 'decreasing',
    clstmethod = 'ssom', ssomt1 = 2, ssomt2 = 2, clusters = 3,
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22, colvec = colvec,
    o_min = c(NA, NA, NA, 4), o_max = c(NA, NA, NA, 16)
)

dev.print(pdf, file='ff.pdf'); dev.print(png, file='ff.png', height=8.27, width=11.7, res=72, units = "in");
```

## [Assembled figure (click for full page)](FigCe10L.html#/all) + [Printable - static HTML](static.html)
<iframe src="FigCe10L.html#/all"></iframe>

#Alternative figures

## [Plot D](fdalt.pdf) - reanked by CpG in preomoter region

```{r  heatmap_cpg_pre, echo=FALSE, message=FALSE, include=FALSE}
temp <- getPlotSetArray(tracks = MotifSetup()$addMotif('GC'),features = features[5], refgenome = 'ce10', bin = 10L, xmin = 100L, xmax = 50L)
ord <- plotHeatmap(temp, sortrows = 'decreasing', clusters = 1)
features[5]  %>% import.bed   %>% .[ord$FinalOrder]  %>% export.bed('test.bed')
M2 <- getPlotSetArray(
    tracks = ms,
    features = 'test.bed',
    refgenome = 'ce10',
    bin = 10L, xmin = 1000L, xmax = 1500L
)
```

```{r heatmap_cpg, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}

plotHeatmap(
    M2, labels = M2$as.array()  %>% colnames %>% strsplit('_') %>% sapply('[[', 1),
    raster = TRUE, include = c(T,F,F,F), sortrows = 'none', clusters = 3, 
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22, colvec = colvec,
    o_min = c(-1, -2, -1, 0), o_max = c(3, 8, 4, 16)
)
dev.print(pdf, file='fdalt.pdf')

```

## [Plot D](fdalt2.pdf) - reanked by CpG w/o clustering

```{r heatmap_cpg_2, message=FALSE, paper = 'a4r', warning=FALSE, echo=TRUE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}

plotHeatmap(
    M2, labels = M2$as.array()  %>% colnames %>% strsplit('_') %>% sapply('[[', 1),
    raster = TRUE, include = c(T,F,F,F), sortrows = 'none', clusters = 0, 
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22, colvec = colvec,
    o_min = c(-1, -2, -1, 0), o_max = c(3, 8, 4, 16)
)
dev.print(pdf, file='fdalt2.pdf')

```


## [Plot D](fdalt3.pdf) - AATAAA TTS (endpoint) motif

```{r heatmap_AATAAA, message=FALSE, paper = 'a4r', warning=FALSE, echo=FALSE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
MotifSetup()$addBigWig(tracks[2])$addBigWig(tracks[3])$addBigWig(tracks[1]) -> ms2
ms2$addMotif('AATTAAA', revcomp = FALSE)
M3 <- getPlotSetArray(
    tracks = ms2,
    features = features[5],
    refgenome = 'ce10',
    bin = 10L, xmin = 1000L, xmax = 1500L, type = 'ef'
)
plotHeatmap(
    M3, labels = M3$as.array()  %>% colnames %>% strsplit('_') %>% sapply('[[', 1),
    raster = TRUE, include = c(F,F,F,T), sortrows = 'decreasing', clusters = 0, 
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22, colvec = colvec,
    o_min = c(-1, -2, -1, 0), o_max = c(3, 8, 4, 2)
)
dev.print(pdf, file='fdalt3.pdf')
```

## [Plot D](fdalt4.pdf) - TTTCAG motif on 3' of exons
```{r heatmap_TTTCAG, echo=FALSE, message=FALSE, paper = 'a4r', warning=FALSE, dev='png', fig.align='center', fig.height=8.27, fig.width=11.7, out.height='450', out.width="auto"}
require(AnnotationDbi)
db <- loadDb( system.file("anno/WBcel235_EnsemblGenes77_TxDb.sqlite", package = 'JADBtools') )
exons(db)  %>% sample(2000) %>% export.bed('exons.bed')
MotifSetup()$addBigWig(tracks[2])$addBigWig(tracks[3])$addBigWig(tracks[1]) -> ms3
ms3$addMotif('TTTCAG', revcomp = FALSE)
M4 <- getPlotSetArray(
    tracks = ms3,
    features = 'exons.bed',
    refgenome = 'ce10',
    bin = 10L, xmin = 1000L, xmax = 1500L, type = 'ef'
)
plotHeatmap(
    M4, labels = M4$as.array()  %>% colnames %>% strsplit('_') %>% sapply('[[', 1),
    raster = TRUE, include = c(F,F,F,T), sortrows = 'decreasing', clusters = 0, 
    cex.legend = 18, cex.main = 36, cex.lab = 28, cex.axis = 22, colvec = colvec,
    o_min = c(-1, -2, -1, 0), o_max = c(3, 8, 4, 2)
)
dev.print(pdf, file='fdalt4.pdf')

```



